---
title: "R Notebook"
output: html_notebook
---

Get free account at the Pew Research Center: http://www.pewresearch.org/download-datasets/

Once you have your account, from the main page (pewresearch.org) go to datasets, and then Internet, Science, and Tech, and download the  Jan. 3-10, 2018 - Core trends survey. (Here's the page it should be on: http://www.pewinternet.org/dataset/jan-3-10-2018-core-trends-survey/)

It should download a directory called 


Then upload the file called "January 3-10, 2018 - Core Trends Survey - CSV.csv" to the Files section of R Studio. Make sure to keep the file called "January 3-10, 2018 - Core Trends Survey - Questionnaire.docx" because it has the list of variables and what they mean.

```{r}
library(tidyverse)
```


```{r}
pew <- read_csv("January 3-10, 2018 - Core Trends Survey - CSV.csv")
```



```{r}
names(pew)
```


These variables are mostly coded as numbers, like sex: Male = 1, Female = 2. But these are better represented as factors in R.

You can refer to a variable in a dataframe with a dollar sign: dataframe$var. So to refer to the sex variable in pew: pew$sex.

Change pew$sex to a factor with factor(pew$sex). Then show the levels with levels(pew$sex).

```{r}
pew$sex <- factor(pew$sex)
levels(pew$sex)


```


We want to change the level names to something more descriptive than 1 and 2. We use fct_recode() to do this. The Questionnaire says 1 = Male and 2 = Female. 

```{r}
pew$sex <- fct_recode(pew$sex, "Male" = "1", "Female" = "2")
levels(pew$sex)
```



Here is another variable, called intfreq, which asked "About how often do you use the internet?" Possible responses are 1: Almost constantly, 2: Several times a day, 3: About once a day, 4: Several times a week, 5: Less often, 8: Don't know, 9: Refused to answer. (Those are listed in the Questionnaire document.)

First, convert it into a factor and look at the levels:

```{r}
pew$intfreq <- factor(pew$intfreq)
levels(pew$intfreq)
```

Here is the code for converting each number to a verbal label. You don't have to split the command onto multiple indented lines like I've done, but it makes it easier to read. There are two missing data numbers: 8 for "Don't know" and 9 for "Refused to answer." Let's just get rid of both of those by setting them to NULL. They will appear as missing data which we can exclude later.

```{r}
pew$intfreq <- fct_recode(pew$intfreq,
                          "Almost constantly" = "1",
                          "Several times/day" = "2",
                          "Once per day" = "3",
                          "Several times/week" = "4",
                          "Less than once/day" = "5",
                          NULL = "8",
                          NULL = "9")
levels(pew$intfreq)
```


A very common way to look at survey data is with crosstabs, i.e., one categorical variable crossed with another. If we want to see how often Males and Females say they use the internet, we can make a table of the counts in each category:

```{r}
pew %>% 
  count(sex, intfreq)
```

To drop the missing data, use drop_na().

```{r}
pew %>% 
  drop_na(intfreq) %>% 
  count(sex, intfreq)
```


A graph would be much easier to read than a table. Take the above code and stick a ggplot bar graph at the end of it.


```{r}
pew %>% 
  drop_na(intfreq) %>% 
  ggplot(aes(x = intfreq, fill = sex)) +
  geom_bar()
```

The viridis color palette is considered by most people to be nicer than the default ggplot colors. Copy-paste the above code chunk below, and:
1. Add the line scale_fill_viridis_d() to the end of the above, and don't forget to add + to the end of the previous line.  
2. Flip it so the bars are horizontal by including the line coord_flip() as well.


```{r}
pew %>% 
  drop_na(intfreq) %>% 
  ggplot(aes(x = intfreq, fill = sex)) +
  geom_bar() +
  scale_fill_viridis_d() +
  coord_flip()
```


Here's another variation on a bar chart that makes it easy to compare the proportions of females and males in each group. 

Copy-paste the above chunk below, and then in the parentheses of geom_bar(), put position = "fill". 

```{r}
pew %>% 
  drop_na(intfreq) %>% 
  ggplot(aes(x = intfreq, fill = sex)) +
  geom_bar(position = "fill") +
  scale_fill_viridis_d() +
  coord_flip()
```


Another variation is better for seeing raw numbers rather than proportions:

1. In the parentheses of geom_bar(), but position = "dodge" to place the bars right next to one another. 
2. I don't like the gray background in ggplots. To give it a white background, include the line theme_minimal().

```{r}
pew %>% 
  drop_na(intfreq) %>% 
  ggplot(aes(x = intfreq, fill = sex)) +
  geom_bar(position = "dodge") +
  scale_fill_viridis_d() +
  coord_flip() +
  theme_minimal()
```



To create side-by-side graphs, one for men and one for women, use facet_wrap(vars(sex)). It creates side-by-side graphs based on whatever variable or variables you put into vars().

In addition, create some better labels with labs(). Include labs(y = "Number of people", x = "How often do you use the internet", title = "Frequency of Internet Usage"). Remember that we flipped the coordinates, so the x and y axes are flipped!
```{r}
pew %>% 
  drop_na(intfreq) %>% 
  ggplot(aes(x = intfreq)) +
  geom_bar() +
  facet_wrap(vars(sex)) +
  coord_flip() +
  theme_minimal() +
  labs(y = "Number of people", x = "How often do you use the internet", title = "Frequency of Internet Usage")
```



Assignment: 

home4nw: Do you currently subscribe to internet service at home? 1 = Yes, 2 = No, 8 = Don't know, 9 = Refused

Do you use the following social media sites:   (1 = Yes, 2 = No)
web1a: Twitter
web1b: Instagram
web1c: Facebook
web1d: Snapchat
web1e: Youtube


How hard would it be for you to give up...  
1 = Very hard, 2 = Somewhat hard, 3 = Not too hard, 4 = Not hard at all, 5 = Impossible, 6 = Do not use, 8 = Don't know, 9 = Refused
pial5a: Television
pial5b: Cell phone
pial5c: Internet
pial5d: Social media


pial11:
Would you say the internet has mostly been a GOOD thing or a BAD thing for society?
1 = Good thing, 2 = Bad thing, 3 = Some of both, 8 = Don't know, 9 = Refused


pial12
How about you, personally? Overall, when you add up all the advantages and 	disadvantages of the internet, would you say the internet has mostly been a GOOD thing or a BAD thing for 	you? 
1 = Good thing, 2 = Bad thing, 3 = Some of both, 8 = Don't know, 9 = Refused


marital
1 = Married, 2 = Living with a partner, 3 = Divorced, 4 = Separated, 5 = Widowed, 6 = Never married, 8 = Don't know, 9 = Refused

educ2
1 = Less than HS, 2 = Some HS, 3 = HS graduate, 4 = Some college, 5 = Associate degree, 6 = College degree, 7 = Some grad school, 8 = Grad degree, 98 = Don't know, 99 = Refused

emplnw
1 = Employed full time, 2 = Employed part-time, 3 = Retired, 4 = Not employed, 5 = Self-employed, 6 = Disabled, 7 = Student, 8 = Other, 98 = Don't know, 99 = Refused